Bayesian exponential random graph modeling of KWL Burial network

### 2020-09-28 analysis/paper/paper.Rmd
add sentences to method section

We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014]. Bergm provides a framework for Bayesian parameter estimation on the result of exponential random graph models, which depends on ergm packages for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008]. Although exponential random graph models can better describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013]. A normalizing constant is a step to make probability distribution to integrate to one, which becomes hard to compute for larger set of networks [@Caimo2017bayesian]. Another problem is the degeneracy that means the probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014]. To overcome these limitations, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the exchange algorithm, which solves computational problems in ERGMs [@Caimo2011].

@@

Markov chain Monte Carlo (MCMC) is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].

>>>>

We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014], which built upon ergm for model specification [@Hunter2008; @Morris2008; @Handcock2008]. Although exponential random graph models can describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013]. A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian]. Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014]. To overcome these limitations in ERGM, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011]. This alleviate above computational problems. By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters. Posterior probability enables us to estimate the effect of ERGM parameter by observing the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian]. The criteria is that the odds ratio grater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.

@@

Markov chain Monte Carlo (MCMC) algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].
